Missing The Point: Non-Convergence in Iterative Imputation Algorithms

### 2020-11-18 Manuscript.Rmd
Add new bullet point intro, combine figures in case study part

@@

@@

```{r echo=FALSE}
# p <- simulation_plot( v ="cov", name = "Coverage rate") + 
#   list(geom_hline(yintercept = 0.95, color = "grey")) +
#   ylim(0.85, 1)

@@

```{r echo=FALSE}
# autocorrelation plots (not used)
# casestudy_plot(v="ac.max.qhat") #+ list(scale_y_continuous(limits = c(0.65,1)))
# casestudy_plot(v="ac.max.lambda")
```

```{r echo=FALSE}
# p <- plot(mids)
# p[1:2,1] # %>% plotly::ggplotly()
param %>% 
  ggplot() + 
  geom_line(aes(x = it, y = mu.hgt, color = as.factor(m))) + 
  theme_classic() + 
  labs(x = "Iteration number",
       y = "Chain mean variable 'hgt'",
       color = "Imputation")
```

From the traceplot of the chain means it seems that mixing improves up-to 10 iterations, while trending is only apparent in the first three iterations.

<!-- and there is some trending up-to 20 iterations as well. The chain variances are more or less stationary after 5 iterations, and mixing requires about 20 iterations. -->

```{r echo=FALSE}
# p <- casestudy_plot(v="est", name = "Regression estimate")
# p %>% plotly::ggplotly()
```

```{r echo=FALSE}
p <- casestudy_plot(v="r.hat.max.mu.hgt", name = "Rhat chain means") + 
  list(
    geom_hline(yintercept = 1.2, color = "grey")#, 
    #scale_y_continuous(limits = c(1,5))
    )
p # %>% plotly::ggplotly()
```

This figure shows that 7 iterations are required before the $\widehat{R}$-values of the chain means drop below the threshold for non-convergence.

```{r echo=FALSE}
p <-
  casestudy_plot(v = "r.hat.max.qhat", name = "Rhat regression estimate") +
  list(geom_hline(yintercept = 1.2, color = "grey")#,
       #scale_y_continuous(limits = c(1, 2.5))
       )
p # %>% plotly::ggplotly()
```

The scientific estimate reaches the threshold much sooner, when $it=14$.

```{r echo=FALSE}
p <- casestudy_plot(v = "r.hat.max.lambda", name = "Rhat lambda") +
  list(geom_hline(yintercept = 1.2, color = "grey")#,
       #scale_y_continuous(limits = c(1, 2.5))
       )
p # %>% plotly::ggplotly()
```

According to the $\widehat{R}$-values of the novel parameter, at least 15 iterations are required.

>>>>

## One

- iterative imputation has become the de facto standard approach to accommodate missing data

- the aim is to draw valid inference: to get unbiased, confidence-valid estimates

- the estimates are obtained by separating the missing data problem from the scientific problem

- the missing data problem is solved by imputing (i.e., filling in) the missing values in the incomplete dataset, using an algorithm such as MICE

- then the analysis of scientific interest can be performed on the completed data, which yields the scientific estimates

- to obtain valid estimates, both the missing data problem and the scientific problem should be solved correctly

- one aspect that is often overlooked is the convergence of the algorithm that is used to generate imputations

## Two

- there is no consensus on the convergence properties of iterative imputation, and there has not been a systematic study on how to evaluate the convergence of these algorithms

- the current practice is visual inspection, but this may not be adequate because 1, 2, 3 (see original text)

- it is difficult because there is not at single point at which convergence is reached, because the algorithm will produce some fluctuation even after convergence (and a ... distribution may not exist because of univariate modeling of multivariate distribution???)

- therefore, we should look at *non*-convergence

- there are non-convergence identifiers for other iterative algorithms (such as MCMC), but it is not known whether these are appropriate for imputation algorithms as well

## Three

- in this paper, we explore different methods for identifying non-convergence in iterative imputation algorithms

- since the ultimate goal is to have inferential validity, we use this as performance measure in a simulation study: if the estimates are unbiased and confidence-valid, the non-convergence is not influential, apparently

- since we typically don't have the ground truth (to calculate bias and coverage), we develop guidelines for interpreting the non-convergence identifiers in practice

@@

## Original text

@@

@@

```{r case, fig.cap = "Case study.", echo=FALSE}
trace <- param %>% 
  ggplot() + 
  geom_line(aes(x = it, y = mu.hgt, color = as.factor(m))) + 
  theme_classic() + 
  labs(x = "Iteration number",
       y = "Chain mean variable 'hgt'",
       color = "Imputation")
means <- casestudy_plot(v="r.hat.max.mu.hgt", name = "Rhat chain means") + 
  list(geom_hline(yintercept = 1.2, color = "grey"))
qhats <-
  casestudy_plot(v = "r.hat.max.qhat", name = "Rhat regression estimate") +
  list(geom_hline(yintercept = 1.2, color = "grey"))
lambdas <- casestudy_plot(v = "r.hat.max.lambda", name = "Rhat lambda") +
  list(geom_hline(yintercept = 1.2, color = "grey"))
(trace + means) / (qhats + lambdas) + plot_layout(guides="collect") + plot_annotation(tag_levels = 'A')

From the traceplot of the chain means (see \@ref(fig:case)A) it seems that mixing improves up-to 10 iterations, while trending is only apparent in the first three iterations.

This figure (\@ref(fig:case)B) shows that 7 iterations are required before the $\widehat{R}$-values of the chain means drop below the threshold for non-convergence.

The scientific estimate reaches the threshold much sooner, when $it=14$ (see \@ref(fig:case)C).

According to the $\widehat{R}$-values of the novel parameter, at least 15 iterations are required (see \@ref(fig:case)D).
